---
title: "DATA 605 - Week 8 Discussion"
author: "Amber Ferger"
date: "3/17/2020"
output: html_document
---

# Chapter 7: Sums of Random Variables, Question 11 (pg.291) 

## Question 
**Assume that you are playing craps with dice that are loaded in the following way: faces two, three, four, and five all come up with the same probability $\frac{1}{6} + r$. Faces one and six come up with probability $\frac{1}{6} − 2r$, with $0 < r < .02$. Write a computer program to find the probability of winning at craps with these dice, and using your program find which values of $r$ make craps a favorable game for the player with these dice.**  

## Solution  
### How to Play Craps

1. Players start by placing a bet. 
2. The sequence starts by the shooter rolling the dice. 
3. Bettors win if the roll is a 7 or 11 and lose if it is a 2, 3, or 12. Any other number becomes the point. 
4. If a point number is rolled, the shooter keeps rolling until either (1) the same number is rolled again (the player wins) or (2) a 7 is rolled (the player loses).

This is what we will use to construct our computer program. 


### Probabilities


```{r, warning=FALSE, echo=FALSE, message=FALSE, results ='hide'}
library(tidyverse)

```

We'll first define the values of $r$ to loop over. We'll also define a list that will hold the probabilities of winning for each value of $r$. 
``` {r}

rStart <- 0
rStop <- 0.02
rBy <- 0.0001
finalProbs <- c()

```


Next, we will define the probability of rolling each sum. We'll also define the probability of winning as the probability of getting a 7 or an 11 on the first roll plus the probability of winning from a point game. 

``` {r}

for (r in seq(rStart, rStop, by=rBy)){

  initialBet <- 2

  # define probabilities for each die roll
  p2 <- (1/6) + r
  p3 <- (1/6) + r
  p4 <- (1/6) + r
  p5 <- (1/6) + r
  
  p1 <- (1/6) - (2*r)
  p6 <- (1/6) - (2*r)
  
  # define probabilities for the sums of the die
  # probabilities for losing
  pRoll2 <- p1*p1
  pRoll3 <- p1*p2 + (p2*p1)
  pRoll12 <- p6*p6
  
  # probabilities for winning
  pRoll7 <- (p1*p6) + (p2*p5) + (p3*p4) + (p5*p2) + (p4*p3) + (p6*p1)
  pRoll11 <- (p5*p6) + (p6*p5)

  # point probabilities
  pRoll4 <- (p1*p3) + (p2*p2) + (p3*p1)
  pRoll5 <- (p1*p4) + (p2*p3) + (p3*p2) + (p4*p1)
  pRoll6 <- (p1*p5) + (p2*p4) + (p3*p3) + (p4*p2) + (p5*p1)
  pRoll8 <- (p2*p6) + (p3*p5) + (p4*p4) + (p5*p3) + (p6*p2)
  pRoll9 <- (p3*p6) + (p4*p5) + (p5*p4) + (p6*p3)
  pRoll10 <- (p4*p6) + (p5*p5) + (p6*p4)
  
  # define the probabilities of winning or losing on the first round
  pLoss <- pRoll2 + pRoll3 + pRoll12
  pWin <- pRoll7 + pRoll11
  
  # define the probabilities of winning after rolling a point number
  pWin4 <- pRoll4 * (pRoll4 / (pRoll4+pRoll7))
  pWin5 <- pRoll5 * (pRoll5 / (pRoll5+pRoll7))
  pWin6 <- pRoll6 * (pRoll6 / (pRoll6+pRoll7))
  pWin8 <- pRoll8 * (pRoll8 / (pRoll8+pRoll7))
  pWin9 <- pRoll9 * (pRoll9 / (pRoll9+pRoll7))
  pWin10 <- pRoll10 * (pRoll10 / (pRoll10+pRoll7))
  
  # define the final probability of winning
  pWinFinal <- pWin + pWin4 + pWin5 + pWin6 + pWin8 + pWin9 + pWin10
  expecVal <- pWin*(initialBet*2) + pLoss*(-1*initialBet) 
  
  finalProbs <- c(finalProbs,c(r,pWinFinal))
  
}

```


Now that our data has been collected, we will restructure our list so that the value of $r$ is in one column and the probability of winning is in the other. 

``` {r}
# every other element
rVals <- seq(1,length(finalProbs),2)
R <- finalProbs[rVals]
WIN <- finalProbs[-rVals]


outcome <- do.call(rbind, Map(data.frame, R=R, WIN=WIN))

```

We can visualize the probability of winning as $r$ increases: 
```{r}
plot(outcome$R,outcome$WIN)
```

From this, we can see that the higher the value of $r$, the more likely we are to win. To calculate the winnings, let's look at what would happen if we bet \$5 that we will win the game. 
```{r}

outcome$EXPEC_VAL <- (outcome$WIN * 5) - ((1-outcome$WIN)*5)
plot(outcome$R, outcome$EXPEC_VAL)

```

It becomes profitable to bet on winning once the value of $r$ is greater than: 

``` {r}

outcome %>%
  filter(EXPEC_VAL > 0) %>%
  select(R) %>%
  summarise(min(R))

```