---
title: "DATA 605 - Week 5 Discussion"
author: "Amber Ferger"
date: "2/25/2020"
output: html_document
---
  
# Chapter 1: Discrete Probability Distributions, Question 6 (pg.13) 

## Question
**In Las Vegas, a roulette wheel has 38 slots numbered 0,00,1,2,...36. The 0 and 00 slots are green and half of the remaining 36 slots are red and half are black. A croupier spins the wheel and throws in an ivory ball. If you bet $1 on red, you win $1 if the ball stops in a red slot and otherwise you lose $1. Write a program to find the total winnings for a player who makes 1,000 bets on red.**  

## Solution  

First, let's define the probability of landing on each color slot: 
```{r}

numG <- 2
total <- 38
numB <- (total-numG)/2
numR <- (total-numG)/2

probG <- numG/total
probB <- numB/total
probR <- numR/total

```

Next, we will define our sample space, probabilities, and initialize the number of bets: 
```{r}
colors <- c('G','B','R')
probs <- c(probG, probB, probR)
numBets <- 1000

```

Now we will use the sample function to sample from our colors list: 
```{r}
slots <- sample(colors,
                size = numBets,
                replace = TRUE,
                prob = probs)
```

Finally, we will compute the total winnings for a player that makes 1,000 bets on red:
```{r}

# we will initialize our finalWinnings variable at $1, as this is the amount the player initially bets
finalWinnings <- 1

for (i in 1:numBets){
  if(slots[i]=='R'){
    finalWinnings <- finalWinnings+1
  } else {
    finalWinnings <- finalWinnings-1
  }
}

```

The player's final winnings are:
```{r}

finalWinnings

```