---
title: "DATA 605 - Week 16 Discussion"
author: "Amber Ferger"
date: "5/11/2020"
output: html_document
---

## Exercises 14.2, Problem 5 (pg. 858)
**Sketch the given vector ﬁeld over the rectangle with opposite corners $(−2,−2)$ and $(2,2)$, sketching one vector for every point with integer coordinates (i.e., at $(0,0),(1,2)$,etc.).**
$$\vec{F} = <x,0>$$

First, let's set up our grid of values. We will expand the grid so that we can graph our vector field in full: 
```{r}

vals <- seq(-2,2)
df <- data.frame(x = integer(),
                 y = integer())

for (x in vals){
  for (y in vals){
    df <- rbind(df,data.frame(x,y))
  }
}

plot(df$x, df$y, main="", xlab="", ylab="", xlim = c(-4,4), ylim = c(-4,4))

```

Now that we have our original $(x,y)$ value pairs, we can calculate our vectors and final coordinate values.

```{r}

df$x_new <- df$x
df$y_new <- 0

df$x_final <- df$x + df$x_new
df$y_final <- df$y + df$y_new

head(df)

```

Finally, we will plot:

```{r, warning = FALSE}

plot(df$x, df$y, main="", xlab="", ylab="", xlim = c(-4,4), ylim = c(-4,4))
s <- seq(nrow(df))
arrows(df$x[s],df$y[s],df$x_final[s],df$y_final[s])

```

From this, we can see that at $x=0$, the resulting vectors all have value $<0,0>$. As we increase the value of $x$, the magnitude of the resulting vector also increases. 