---
title: "DATA 605 - Week 7 Discussion"
author: "Amber Ferger"
date: "3/11/2020"
output: html_document
---

# Chapter 5: Distributions and Densities, Section 1- Important Distributions, Question 18 (pg.199)
## Question  
**A baker blends 600 raisins and 400 chocolate chips into a dough mix and, from this, makes 500 cookies.**  

**a. Find the probability that a randomly picked cookie will have *no* raisins.**  
**b. Find the probability that a randomly picked cookie will have *exactly two* chocolate chips.**  
**c. Find the probability that a randomly chosen cookie will have *at least two bits* (raisins or chips) in it.**  

## Solution  
Let's first create variables for our raisins, chips, and cookies:
```{r}
numR <- 600
numCC <- 400
numC <- 500

avgRperC <- numR / numC
avgCCperC <- numCC / numC
avgBitsperC <- (numR + numCC) / numC

```

### a. Find the probability that a randomly picked cookie will have no raisins.   

We know that the average number of raisins per cookie is: 
```{r}
avgRperC
```

As a gut check, we would expect the probability of no raisins per cookie to be low. We will model this using a Poission Distribution: 
$$P(X=x)=\frac{e^{-\lambda}\lambda^{x}}{x!}$$  

where $\lambda$ is the average number of raisins per cookie, and $x$ is 0 raisins. 
```{r}
x <- 0
noRaisins <- (exp(-avgRperC)* avgRperC^x)/factorial(x)
noRaisins

```


### b. Find the probability that a randomly picked cookie will have exactly two chocolate chips.    
We know that the average number of chocolate chips per cookie is: 
```{r}
avgCCperC
```

As a gut check, we know that the probability of exactly 2 chocolate chips should be low. Once again, we will model this using a poisson distribution, this time with $\lambda$ as the average number of chocolate chips per cookie, and $x$ as 2 chips.
```{r}

x2 <- 2
twoChips <- (exp(-avgCCperC)* avgCCperC^x2)/factorial(x2)
twoChips

```

### c. Find the probability that a randomly chosen cookie will have at least two bits (raisins or chips) in it.  

We know that the average number of bits per cookie is: 
```{r}
avgBitsperC
```

We will model this by calculating the complement of the probability of there being 0 or 1 bits in the cookie:  
$$P(X>2)=1-[P(X=0)+P(X=1)]$$
```{r}

x3 <- 0
x4 <- 1
prob0 <- (exp(-avgBitsperC)* avgBitsperC^x3)/factorial(x3)
prob1 <- (exp(-avgBitsperC)* avgBitsperC^x4)/factorial(x4)

prob2orMore <- 1 - (prob0+prob1)

prob2orMore

```
